A Computational Account of Default-Mode Function\\by Control Theory and Reinforcement Learning

### 2016-07-16 paper.tex
impose structure

\section{Introduction to the Default-Mode Network}
%
\paragraph{The Human Default Mode Network.}
scene construction processes when envisioning past, ficti- tious, and future events (Hassabis et al., 2007; Spreng et al., 2009).

@@

\paragraph{The Predictive Coding Framework.}
\begin{itemize}
  \item It is one of the most frequently cited hypotheses of default mode function
  \item Within this model the brain is conceptualized as a Bayesian machine
  \item Predictive coding is a framework that, in a hierarchical setting,
  is equivalent to empirical Bayesian inference
  \item External sensory input is feed-forward/bottom-up processed in the brain
  \item Their processing is compared against expected input
  \item The prediction error, i.e. the difference between sensory input
  and internal prediction, is computed at each level and passed to
  higher levels via forward connections
  \item in case of a mismatch, back-projections top-down modulate
  input processing
  \item ‘contextual integration’ for top-down modulation of sensorimotor
  processing by context-specific a-priori information.
  \item Closely related to Friston's free-energy principle: brain as inference engine
  biological systems, including brains, must minimize the long-term average of surprise;
  formulate perception as a constructive process based on internal or generative models
  \item A generative model of the world
\end{itemize}

\paragraph{Control Theory.}
\begin{itemize}
\item mathematically formalized dynamics systems with input, output and
feedback loops that produce error signals for adaptation
\item typically realized by (non-linear) differential equations
\item has been used to describe neural systems in the past
\item gradient update = learn/adapt system dynamics by error occurrence
\item related to notions of game theory
\item switching between states -> typically reduces to a sequential processing model
\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
\item policy: action to take given a state
\item value: prediction of future reward
\item model: predict what the environment will do next
\item model-baseed RI learning: we try to understand the environment; it is initially unknown
\item special to RL: exploitation vs. exploration
\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
\end{itemize}

\paragraph{States and Actions in Neuroscience}

state-dependent variables such as hunger, thirst and
even wealth

\paragraph{Anatomy}
 involvement of the PMC across the boundaries of classical cognitive domains
mPFC and PCC = highest metabolic consumption in the brain/highest level of basal glucose energy consumption in humans

@@

@@

\item
only cluster 2 was congruently coupled (across MACM and RSFC) with the amygdala (involved in significance evaluation) and the nucleus accumbens (involved in reward evaluation)
s well as exclusively functionally asso- ciated with facial appraisal. Similarly, the human vPCC was the only PMC region connected to the laterobasal (rather than centromedial or superficial) nuclei group of the amygdala (Bzdok et al., 2013b), which is an amygdalar subregion probably devoted to continuously scanning environmental input for biological significance (Adolphs, 2010; Aggleton et al., 1980; Bzdok et al., 2011; Ghods-Sharifi et al., 2009; LeDoux, 2007).
\end{itemize}

\item
\item Bálint's syndrome: neurological disorder of conscious
awareness resulting from damage in bilateral parietal cortex;
"psychic paralysis of gaze", that is, the inability for internally
motivated, targeted saccades hindering fixation of currently not
attended features in the field of view (i.e., oculomotor apraxia).
unable to bind various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as unable to navigate hand movement to a targeted
object helped by vision (i.e., optic ataxia).
difficulty avoiding running into objects placed ahead;
challenged in estimating distances as well as perceiving
spatial depth, object sizes and object orientation

  \item
    vPCC:
the ventral visual stream—the “what” system for object process- ing (Ungerleider and Haxby, 1994; Ungerleider and Mishkin, 1982; Vogt et al., 2006), contrary to the dPCC's conceivable relation to the dorsal “where” system
\item
  dPCC:
This view is in line with the well-known integration of the dPCC in the dorsal visual stream—the “where” system for spatial processing (Ungerleider and Haxby, 1994; Ungerleider and Mishkin, 1982; Vogt et al., 2006), in contrast to vPCC's relation to the ventral “what” system

\item
RSC:
Regarding memory processing, lesions in the RSC, as compared to other PMC regions, is most consistently associated with anterograde and retrograde memory impairments of different types of sensory infor- mation in rabbits (Gabriel and Talk, 2001) and humans (Rudge and Warrington, 1991; Valenstein et al., 1987).
current findings and earlier lesion reports can be parsimoniously reconciled by the previously proposed notion that the RSC mediates between the organism's egocentric (i.e., focused on sensory input) and allocentric (i.e., focused on world knowledge) perspective frames (Burgess, 2008; Epstein, 2008; Valiquette and McNamara, 2007).

\item
BOTH TPJ:

RTPJ:
\item
Corbetta2008:
Although both attentional networks have been most exten- sively investigated in vision, the available evidence indicates a supramodal function (Driver and Spence, 1998; Macaluso et al., 2002). The ventral network (right TPJ, right IFG) registers salient events in the environment not only in the visual but also in the auditory and tactile modalities (Downar et al., 2000)
Several lines of evidence indicate that two cortico-cortical neural systems are involved in attending to environmental stimuli (Corbetta and Shulman, 2002). A dorsal frontoparietal network. A second system, the ventral frontoparietal network, is not ac- tivated by expectations or task preparation but responds along with the dorsal network when behaviorally relevant objects (or targets) are detected (Corbetta et al., 2000). Both dorsal and ventral networks are also activated during reorienting, with en- hanced responses during the detection of targets that appear at unattended locations. For example, enhanced responses are observed when subjects are cued to expect a target at one location but it unexpectedly appears at another (i.e., ‘‘invalid’’ targets in the Posner spatial cueing paradigm)
or when a target appears infre- quently, as in ‘‘oddball’’ paradigms.

@@

qualify RTPJ as a po- tential switch between exteroceptive and interoceptive mind sets implemented by their networks. The reciprocal nature of those mind sets is well illustrated by consistent reports of continuous shifting be- tween externally (i.e., sensory) and internally (i.e., self- and presumably social/memory-) oriented processing in controlled laboratory settings and daily routine (Smallwood et al., 2007).

\item LTPJ:

\item
the difference between the two task groups could be parsimoniously explained by the required atten- tion to either the external world or self.
world -> policy matrix driven computation
self -> value matrix driven computation

@@

LTPJ:

@@

\item a lesion study on disturbed sleep (i.e., a state of mind
independent of sensory stimulation but dependent on
internally generated information) exclusively identified
the dmPFC (Koenigs et al., 2010). Third, another VLSM study exclusively
related the IFG and TPJ, both more strongly connected to the
dmPFC in our study, to inner speech

\item PMC:
  \begin{itemize}
    \item
covert reallocation of spatial attention (Gitelman et al., 1999), mediation between internal and external focus (Leech and Sharp, 2014), computation of environmental statistics (Pearson et al., 2009), and self-referential visuospatial imagery (Cavanna and Trimble, 2006)
  \end{itemize}

@@

@@

\section{Experimental Results}
\paragraph{Serial versus parallel structure discovery and classification.}

\section{Discussion and Conclusion}

>>>>

\section{Introduction to the Default-Mode Network}
%
scene construction processes when envisioning past, ficti- tious, and future events (Hassabis et al., 2007; Spreng et al., 2009).

@@

The Predictive Coding Framework.
\begin{itemize}
  \item It is one of the most frequently cited hypotheses of default mode function
  \item Within this model the brain is conceptualized as a Bayesian machine
  \item Predictive coding is a framework that, in a hierarchical setting,
  is equivalent to empirical Bayesian inference
  \item External sensory input is feed-forward/bottom-up processed in the brain
  \item Their processing is compared against expected input
  \item The prediction error, i.e. the difference between sensory input
  and internal prediction, is computed at each level and passed to
  higher levels via forward connections
  \item in case of a mismatch, back-projections top-down modulate
  input processing
  \item ‘contextual integration’ for top-down modulation of sensorimotor
  processing by context-specific a-priori information.
  \item Closely related to Friston's free-energy principle: brain as inference engine
  biological systems, including brains, must minimize the long-term average of surprise;
  formulate perception as a constructive process based on internal or generative models
  \item A generative model of the world
\end{itemize}

\section{Introduction to Control Theory and Reinforcement Learning}
\paragraph{Control Theory.}
\begin{itemize}
\item mathematically formalized dynamics systems with input, output and
feedback loops that produce error signals for adaptation
\item typically realized by (non-linear) differential equations
\item has been used to describe neural systems in the past
\item gradient update = learn/adapt system dynamics by error occurrence
\item related to notions of game theory
\item switching between states -> typically reduces to a sequential processing model
\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
\item policy: action to take given a state
\item value: prediction of future reward
\item model: predict what the environment will do next
\item model-baseed RI learning: we try to understand the environment; it is initially unknown
\item special to RL: exploitation vs. exploration
\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
\end{itemize}

\section{The Neurobiological Properties of the Default-Mode Network}
 involvement of the PMC across the boundaries of classical cognitive domains
mPFC and PCC = highest metabolic consumption in the brain/highest level of basal glucose energy consumption in humans

@@

\item
covert reallocation of spatial attention (Gitelman et al., 1999), mediation between internal and external focus (Leech and Sharp, 2014), computation of environmental statistics (Pearson et al., 2009), and self-referential visuospatial imagery (Cavanna and Trimble, 2006)

@@

\item
only cluster 2 was congruently coupled (across MACM and RSFC) with the amygdala (involved in significance evaluation) and the nucleus accumbens (involved in reward evaluation)
s well as exclusively functionally asso- ciated with facial appraisal. Similarly, the human vPCC was the only PMC region connected to the laterobasal (rather than centromedial or superficial) nuclei group of the amygdala (Bzdok et al., 2013b), which is an amygdalar subregion probably devoted to continuously scanning environmental input for biological significance (Adolphs, 2010; Aggleton et al., 1980; Bzdok et al., 2011; Ghods-Sharifi et al., 2009; LeDoux, 2007).

\item Bálint's syndrome: neurological disorder of conscious
awareness resulting from damage in bilateral parietal cortex;
"psychic paralysis of gaze", that is, the inability for internally
motivated, targeted saccades hindering fixation of currently not
attended features in the field of view (i.e., oculomotor apraxia).
unable to bind various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as unable to navigate hand movement to a targeted
object helped by vision (i.e., optic ataxia).
difficulty avoiding running into objects placed ahead;
challenged in estimating distances as well as perceiving
spatial depth, object sizes and object orientation

  \item
    vPCC:
the ventral visual stream—the “what” system for object process- ing (Ungerleider and Haxby, 1994; Ungerleider and Mishkin, 1982; Vogt et al., 2006), contrary to the dPCC's conceivable relation to the dorsal “where” system
\item
  dPCC:
This view is in line with the well-known integration of the dPCC in the dorsal visual stream—the “where” system for spatial processing (Ungerleider and Haxby, 1994; Ungerleider and Mishkin, 1982; Vogt et al., 2006), in contrast to vPCC's relation to the ventral “what” system
\end{itemize}

\item RSC:
  \begin{itemize}
\item
Regarding memory processing, lesions in the RSC, as compared to other PMC regions, is most consistently associated with anterograde and retrograde memory impairments of different types of sensory infor- mation in rabbits (Gabriel and Talk, 2001) and humans (Rudge and Warrington, 1991; Valenstein et al., 1987).
current findings and earlier lesion reports can be parsimoniously reconciled by the previously proposed notion that the RSC mediates between the organism's egocentric (i.e., focused on sensory input) and allocentric (i.e., focused on world knowledge) perspective frames (Burgess, 2008; Epstein, 2008; Valiquette and McNamara, 2007).

\item
Neuroscientific investigations on hemispheric specialization have converged to the left cerebral hemisphere as dominant for language functions (Broca 1865, Wernicke 1874, Lichtheim 1885) and the right hemisphere as dominant for attention functions (Gazzaniga, Bogen et al. 1965, Sperry 1982, Stephan, Marshall et al. 2003). 
  \end{itemize}

  \item
    RTPJ:
    \begin{itemize}
\item
Corbetta2008:
Although both attentional networks have been most exten- sively investigated in vision, the available evidence indicates a supramodal function (Driver and Spence, 1998; Macaluso et al., 2002). The ventral network (right TPJ, right IFG) registers salient events in the environment not only in the visual but also in the auditory and tactile modalities (Downar et al., 2000)
Several lines of evidence indicate that two cortico-cortical neural systems are involved in attending to environmental stimuli (Corbetta and Shulman, 2002). A dorsal frontoparietal network. A second system, the ventral frontoparietal network, is not ac- tivated by expectations or task preparation but responds along with the dorsal network when behaviorally relevant objects (or targets) are detected (Corbetta et al., 2000). Both dorsal and ventral networks are also activated during reorienting, with en- hanced responses during the detection of targets that appear at unattended locations. For example, enhanced responses are observed when subjects are cued to expect a target at one location but it unexpectedly appears at another (i.e., ‘‘invalid’’ targets in the Posner spatial cueing paradigm)
or when a target appears infre- quently, as in ‘‘oddball’’ paradigms.

@@

qualify RTPJ as a po- tential switch between exteroceptive and interoceptive mind sets implemented by their networks. The reciprocal nature of those mind sets is well illustrated by consistent reports of continuous shifting be- tween externally (i.e., sensory) and internally (i.e., self- and presumably social/memory-) oriented processing in controlled laboratory settings and daily routine (Smallwood et al., 2007).
\end{itemize}

\item
  LTPJ
  \begin{itemize}
\item
the difference between the two task groups could be parsimoniously explained by the required atten- tion to either the external world or self.
world -> policy matrix driven computation
self -> value matrix driven computation

@@

\item
complex conceptual linguistic functions are proposed to be more strongly left-lateralized -> in line with overall left-lateralization of human language

\item
Episodic and explicit memory retrieval strongly draws on complex semantic processing and contributes to social cognitive processes (see section 4.4.). Syntactic processing, on the other hand, is a core language process that is closely intermingled with semantic processing. It refers to the hierarchical sequencing of words and their meanings (Price, 2010) and is mandatory for sentence processing in both social cognitive and language tasks.

@@

\item a lesion study on disturbed sleep (i.e., a state of mind
independent of sensory stimulation but dependent on
internally generated information) exclusively identified
the dmPFC (Koenigs et al., 2010). Third, another VLSM study exclusively
related the IFG and TPJ, both more strongly connected to the
dmPFC in our study, to inner speech
\end{itemize}

@@

  \end{itemize}
\end{itemize}

@@

\section{Relation to Other Cognitive Models (Danilo)}
\paragraph{Predictive Coding & Bayesian Brain Hypothesis}

\paragraph{Semantic Hypothesis}

\paragraph{Sentinel Hypothesis}

\section{Relation to Other Statistical Models (Elvis)}
\paragraph{Control Theory}

\paragraph{Q-Learning}

\paragraph{Free Energy Principle}

\section{Conclusion}
